---
title: "All categories"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 4
runtime: shiny
resource_files:
- Data/Shapefiles/County shapefiles/MNCounties_MNDOT.cpg
- Data/Shapefiles/County shapefiles/MNCounties_MNDOT.dbf
- Data/Shapefiles/County shapefiles/MNCounties_MNDOT.prj
- Data/Shapefiles/County shapefiles/MNCounties_MNDOT.sbn
- Data/Shapefiles/County shapefiles/MNCounties_MNDOT.sbx
- Data/Shapefiles/County shapefiles/MNCounties_MNDOT.shp.xml
- Data/Shapefiles/County shapefiles/MNCounties_MNDOT.shx
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(tidyverse)
library(sf)
library(ggrepel)
library(scales)
library(shiny)
library(shinycssloaders)
library(ggiraph)
library(kableExtra)
library(rmapshaper)
library(cowplot)
library(DT)
library(htmlwidgets)
library(RColorBrewer)
library(readxl)
library(janitor)
library(lubridate)
library(systemfonts)
reset_font_cache()
library(ggtext)
library(gmodels)

```

```{r loading jon docs and shapefiles, cache=TRUE, include=FALSE}
theme_bar <- theme_bw() +
  theme(panel.grid.major = element_line(color = "grey70", size = 0.1),
        panel.grid.minor = element_blank(),
        axis.ticks = element_blank(),
        axis.text = element_text(face = "bold"),
        panel.border = element_blank(),
        legend.background = element_rect(fill = "transparent", color = "transparent"),
        legend.key = element_rect(fill = "transparent"),
        legend.key.size = unit(1, "lines"),
        legend.margin = margin(0,0,0,0),
        legend.title = element_blank(),
        legend.text = element_text(margin = margin(l = 2)),
        text = element_text(family = "Arial") ,
        plot.title.position = "plot",
        plot.title = element_text(face = "bold"))

theme_line <- theme_bw() +
  theme(legend.background = element_rect(fill = "transparent", color = "transparent"),
        legend.key = element_rect(fill = "transparent"),
        legend.text = element_text(margin = margin(l = 2)),
        panel.grid.minor = element_blank(),
        panel.grid.major = element_line(color = "grey70", size = 0.1),
        axis.ticks = element_blank(),
        axis.text = element_text(face = "bold"),
        panel.border = element_blank(),
        legend.margin = margin(0,0,0,0),
        legend.key.size = unit(1, "lines"),
        text = element_text(family = "Arial") ,
        plot.title.position = "plot",
        plot.title = element_text(face = "bold"))


theme_sf <- theme_bw() +
  theme(axis.text.x=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks=element_blank(),
        panel.background = element_blank(),
        panel.grid.major = element_line(color = "white"),
        panel.border = element_blank(),
        legend.title = element_blank(),
        legend.text = element_text(margin = margin(l = 2)),
        legend.margin = margin(0,0,0,0),
        legend.key.size = unit(1, "lines"),
        text = element_text(family = "Arial") ,
        plot.title.position = "plot",
        plot.title = element_text(face = "bold"))

regions <- read_csv("Data/Join docs/county_regions.csv") %>%
    select(5,6) %>%
    unique() %>%
    mutate(edr = str_replace(edr, "  ", " "),
           planning.region = str_replace(planning.region, " Minnesota", ""),
           planning.region = fct_relevel(planning.region, "Northwest", "Northeast", "Central", "Seven County Mpls-St Paul", "Southwest", "Southeast"),
           edr = fct_relevel(edr, "EDR 1 - Northwest", "EDR 2 - Headwaters", "EDR 3 - Arrowhead", "EDR 4 - West Central", "EDR 5 - North Central", "EDR 6E- Southwest Central", "EDR 6W- Upper Minnesota Valley", "EDR 7E- East Central", "EDR 7W- Central", "EDR 8 - Southwest", "EDR 9 - South Central", "EDR 10 - Southeast", "EDR 11 - 7 County Twin Cities", "Minnesota"))

counties.regions <- read_csv("Data/Join docs/county_regions.csv") %>%
  rename(mif = `MIF Region`) %>%
  mutate(countyfp = formatC(countyfp, width = 3, flag = "0"),
         Name = str_to_title(Name),
         Name = str_replace(Name, "Q", "q"),
         Name = str_replace(Name, "Of The", "of the"),
         Name = str_replace(Name, "Mcleod", "McLeod"),
         Dem_Desc = ifelse(Name == "Minnesota", "Minnesota", Dem_Desc) ,
         edr = str_replace(edr, "  ", " "),
         planning.region = str_replace(planning.region, " Minnesota", ""),
         planning.region = fct_relevel(planning.region, "Northwest", "Northeast", "Central", "Seven County Mpls-St Paul", "Southwest", "Southeast"),
         edr = fct_relevel(edr, "EDR 1 - Northwest", "EDR 2 - Headwaters", "EDR 3 - Arrowhead", "EDR 4 - West Central", "EDR 5 - North Central", "EDR 6E- Southwest Central", "EDR 6W- Upper Minnesota Valley", "EDR 7E- East Central", "EDR 7W- Central", "EDR 8 - Southwest", "EDR 9 - South Central", "EDR 10 - Southeast", "EDR 11 - 7 County Twin Cities", "Minnesota"),
         mif = ifelse(is.na(mif), "TC", mif),
         mif = as.factor(mif),
         mif = fct_relevel(mif, "NW", "NE", "WC", "EC", "SW", "SE", "TC"))


color.ruca <- c("Entirely rural" = "#009933", "Town/rural mix" = "#99CC33", "Urban/town/rural mix" = "#CC9966", "Entirely urban" = "#754C29", "Minnesota" = "black")

color.pr <- c("Northwest" = 	"#4575b4", "Northeast" = "grey", "Central" = "#fee090", "Seven County Mpls-St Paul" = "#d73027", "Southwest" = "#91bfdb", "Southeast" = "#fc8d59", "Minnesota" = "black")

color.edr <- c("EDR 1 - Northwest" = "#b3cde3", "EDR 2 - Headwaters" = "#8c96c6", "EDR 3 - Arrowhead" = "#fe9929", "EDR 4 - West Central" = "#8856a7", "EDR 5 - North Central" = "#810f7c", "EDR 6E- Southwest Central" = "#e5f5f9", "EDR 6W- Upper Minnesota Valley" = "#bdc9e1", "EDR 7E- East Central" = "#99d8c9", "EDR 7W- Central" = "#2ca25f", "EDR 8 - Southwest" = "#74a9cf", "EDR 9 - South Central" = "#0570b0", "EDR 10 - Southeast" = "#d7301f", "EDR 11 - 7 County Twin Cities" = "#d8b365", "Minnesota" = "black")

color.pr.edr <- c ("Northwest" = "#4575b4","Northeast" = "#e0f3f8", "Central" = "#fee090", "Seven County Mpls-St Paul" = "#d73027", "Southwest" = "#91bfdb", "Southeast" = "#fc8d59", "Minnesota" = "black", "EDR 1 - Northwest" = "#b3cde3", "EDR 2 - Headwaters" = "#8c96c6", "EDR 3 - Arrowhead" = "#fe9929", "EDR 4 - West Central" = "#8856a7", "EDR 5 - North Central" = "#810f7c", "EDR 6E- Southwest Central" = "#e5f5f9", "EDR 6W- Upper Minnesota Valley" = "#bdc9e1", "EDR 7E- East Central" = "#99d8c9", "EDR 7W- Central" = "#2ca25f", "EDR 8 - Southwest" = "#74a9cf", "EDR 9 - South Central" = "#0570b0", "EDR 10 - Southeast" = "#d7301f", "EDR 11 - 7 County Twin Cities" = "#d8b365")

mn_counties <- st_read("Data/Shapefiles/county shapefiles/MNCounties_MNDOT.shp", quiet = TRUE) %>%
  ms_simplify(keep = .01, keep_shapes = TRUE) %>%
  rename(countyfp = FIPS_CODE)

```


```{r master basic revenue}
original <- read_csv("Data/Formula/Master.csv") %>%
  mutate(Dem_Desc = fct_relevel(Dem_Desc, "Entirely rural", "Town/rural mix", "Urban/town/rural mix", "Entirely urban"),
         planning.region = fct_relevel(planning.region, "Northwest", "Northeast", "Central", "Seven County Mpls-St Paul", "Southwest", "Southeast"),
         edr = fct_relevel(edr, "EDR 1 - Northwest", "EDR 2 - Headwaters", "EDR 3 - Arrowhead", "EDR 4 - West Central", "EDR 5 - North Central", "EDR 6E- Southwest Central", "EDR 6W- Upper Minnesota Valley", "EDR 7E- East Central", "EDR 7W- Central", "EDR 8 - Southwest", "EDR 9 - South Central", "EDR 10 - Southeast", "EDR 11 - 7 County Twin Cities", "Minnesota"))

group.codes <- read_xlsx("Data/Formula/group codes.xlsx")

var.codes <- read_xlsx("Data/Formula/primary variable codes.xlsx")

categories <- c("bsrev", "exdayr", "gtrev", "smlrev", "decrev", "locaid", "afdc22", "totlep", "totspar", "cexaid", "tsparrev", "eqaid", "hhaid", "penrev")

```

<br>

Below is an analysis of all the categories to determine which of them makes up the largest proportions of total basic education revenue for each school district, as well as schools within school types (charter vs. public), RUCA categories and regions.

<br>

# Proportion of total revenue by category{.tabset}

First let's check to see what categories contribute the most revenue to school districts. Are there categories that provide more than another depending on the type of school, RUCA categories or regions?

First, let's lump all schools together and see what which categories provide the most. As the chart below shows, the basic education revenue provides BY FAR the most to school districts in Minnesota. Very far behind is compensatory which provides 6.4% of the total revenue given to schools.

<br>

```{r categories all school}
cat.all.schools <- original %>%
  filter(group %in% c(1,4)) %>%
  summarize(bsrev = sum(bsrev),
            exdayr = sum(exdayr),
            gtrev = sum(gtrev), 
            smlrev = sum(smlrev), 
            decrev = sum(decrev), 
            locaid = sum(locaid), 
            afdc22 = sum(afdc22), 
            totlep = sum(totlep), 
            totspar = sum(totspar), 
            cexaid = sum(cexaid, na.rm = TRUE), 
            tsparrev = sum(tsparrev), 
            eqaid = sum(eqaid), 
            hhaid = sum(hhaid), 
            penrev = sum(penrev)) %>%
  gather(key = "categories", value = "revenue", 1:14) %>%
  mutate(pct.revenue = revenue / sum(revenue),
         data_id = as.character(seq(n()))) %>%
  left_join(var.codes, by = c("categories" = "Code"))

cat.all.schools.plot <- ggplot(cat.all.schools, aes(fct_reorder(Category, pct.revenue), pct.revenue)) +
  geom_col_interactive(position = "dodge", aes(data_id = data_id, tooltip = paste(Category, "\nTotal revenue: ", dollar(revenue), "\nPercent of total revenue: ", percent(pct.revenue, accuracy = .1), sep = ""))) +
  geom_label(aes(label = percent(pct.revenue, accuracy = .1)), show.legend = FALSE, color = "black", size = 5, position = position_dodge(width = .9)) +
  coord_flip() +
  labs(x="", y = "", color="", title = "Percent share of total general education revenue\nformula")+
  scale_y_continuous(labels=scales::percent)+
  theme_bar+
  theme(legend.position = "none",
        text = element_text(size = 18))


girafe(ggobj = cat.all.schools.plot, width_svg = 10, height_svg = 10) %>%
  girafe_options(opts_selection(type = "none"),
                 opts_sizing(rescale = FALSE))      

```

<br>

Next, let's see if there are any big differences between public and charter schools. The chart below provides the proportion of revenue from each category by type of schools. Charter schools receive a significant larger amount of their revenue from compensatory, english learners, small schools revenue.

<br>

```{r categories by group}
cat.group <- original %>%
  filter(group %in% c(1,4)) %>%
  group_by(group) %>%
  summarize(bsrev = sum(bsrev),
            exdayr = sum(exdayr),
            gtrev = sum(gtrev), 
            smlrev = sum(smlrev), 
            decrev = sum(decrev), 
            locaid = sum(locaid), 
            afdc22 = sum(afdc22), 
            totlep = sum(totlep), 
            totspar = sum(totspar), 
            cexaid = sum(cexaid, na.rm = TRUE), 
            tsparrev = sum(tsparrev), 
            eqaid = sum(eqaid), 
            hhaid = sum(hhaid), 
            penrev = sum(penrev)) %>%
  gather(key = "categories", value = "revenue", 2:15) %>%
  group_by(group) %>%
  mutate(pct.revenue = revenue / sum(revenue)) %>%
  ungroup() %>%
  mutate(data_id = as.character(seq(n()))) %>%
  left_join(var.codes, by = c("categories" = "Code")) %>%
  left_join(group.codes, by = c("group" = "Code")) %>%
  select(Definition, Category, revenue, pct.revenue, data_id)


cat.group.plot <- ggplot(cat.group, aes(fct_reorder(Category, pct.revenue), pct.revenue)) +
  facet_wrap(~Definition) +
  geom_col_interactive(position = "dodge", aes(data_id = data_id, tooltip = paste("School type: ", Definition, "n", Category, "\nTotal revenue: ", dollar(revenue), "\nPercent of total revenue: ", percent(pct.revenue, accuracy = .1), sep = ""))) +
  geom_label(aes(label = percent(pct.revenue, accuracy = .1)), show.legend = FALSE, color = "black", size = 5, position = position_dodge(width = .9)) +
  coord_flip() +
  labs(x="", y = "", color="", title = "Percent share of total general education revenue\nformula")+
  scale_y_continuous(labels=scales::percent)+
  theme_bar+
  theme(legend.position = "none",
        text = element_text(size = 18))


girafe(ggobj = cat.group.plot, width_svg = 10, height_svg = 10) %>%
  girafe_options(opts_selection(type = "none"),
                 opts_sizing(rescale = FALSE))      


```

<br>

Okay, let's see what kind of differences there are by RUCA categories and regions.

The RUCA tables show that as a county group becomes more urban, the percentage of revenue coming from the basic education component decreases. This is because rural areas get more revenue from some of the other components. The components that seem to benefit rural schools more than urban are;

* Declining enrollment
* Equity aid
* Local optional aid
* Sparsity
* Transportation sparsity
* Small schools

There are a few that benefit urban schools more;

* Operating capital aid
* English learners
* Pensions adjustment
* Extended time

It's worth mentioning that compensatory revenue is a significant percentage across all RUCA categories. The highest is in the entirely urban areas, but is quite similar to our entirely rural areas.

When looking at things by geography, there are a few patterns. We will go through each category.

* Basic revenue: the highest percentages are in Central, seven county metro, and Southeast. Most of the EDRs in these regions have well over 80% of their total revenue from this category.
* Compensatory: The highest compensatory percentages exist in EDR 2 - Headwaters, the twin cities, and the north central lakes regions. 
* Declining enrollment: the highest percentage is in Southeast regions.
* English learners: the highest percentages are in Southwest region as well as seven county metro.
* Equity aid: higherst percentages in Northwest and West Central regions.
* Extended time: highest percentages exist in central Minnesota and seven county metro.
* Gifted and talented: even across all of Minnesota.
* Local optional aid: Local optional doesn't have a pattern that I can see.


<br>

## RUCA

<br>

```{r categories by ruca}
cat.ruca <- original %>%
  filter(group %in% c(1,4)) %>%
  drop_na(Dem_Desc) %>%
  group_by(Dem_Desc) %>%
  summarize(bsrev = sum(bsrev),
            exdayr = sum(exdayr),
            gtrev = sum(gtrev), 
            smlrev = sum(smlrev), 
            decrev = sum(decrev), 
            locaid = sum(locaid), 
            afdc22 = sum(afdc22), 
            totlep = sum(totlep), 
            totspar = sum(totspar), 
            cexaid = sum(cexaid, na.rm = TRUE), 
            tsparrev = sum(tsparrev), 
            eqaid = sum(eqaid), 
            hhaid = sum(hhaid), 
            penrev = sum(penrev)) %>%
  gather(key = "categories", value = "revenue", 2:15) %>%
  group_by(Dem_Desc) %>%
  mutate(pct.revenue = revenue / sum(revenue)) %>%
  ungroup() %>%
  mutate(data_id = as.character(seq(n()))) %>%
  left_join(var.codes, by = c("categories" = "Code")) %>%
  select(Dem_Desc, Category, pct.revenue) %>%
  spread(Dem_Desc, pct.revenue)

datatable(cat.ruca, class = "cell-border stripe", filter = "top", rownames = FALSE,
          options = list(columnDefs = list(list(className = "dt-center", targets = 1:4)))) %>%
  formatPercentage(2:5, digits = 1)


```



<br>

## Planning Region

<br>

```{r categories by pr}
cat.pr <- original %>%
  filter(group %in% c(1,4)) %>%
  drop_na(planning.region) %>%
  group_by(planning.region) %>%
  summarize(bsrev = sum(bsrev),
            exdayr = sum(exdayr),
            gtrev = sum(gtrev), 
            smlrev = sum(smlrev), 
            decrev = sum(decrev), 
            locaid = sum(locaid), 
            afdc22 = sum(afdc22), 
            totlep = sum(totlep), 
            totspar = sum(totspar), 
            cexaid = sum(cexaid, na.rm = TRUE), 
            tsparrev = sum(tsparrev), 
            eqaid = sum(eqaid), 
            hhaid = sum(hhaid), 
            penrev = sum(penrev)) %>%
  gather(key = "categories", value = "revenue", 2:15) %>%
  group_by(planning.region) %>%
  mutate(pct.revenue = revenue / sum(revenue)) %>%
  ungroup() %>%
  mutate(data_id = as.character(seq(n()))) %>%
  left_join(var.codes, by = c("categories" = "Code")) %>%
  select(planning.region, Category, pct.revenue) %>%
  spread(planning.region, pct.revenue)

datatable(cat.pr, class = "cell-border stripe", filter = "top", rownames = FALSE,
          options = list(columnDefs = list(list(className = "dt-center", targets = 1:6)),
                         scrollX = TRUE)) %>%
  formatPercentage(2:7, digits = 1)


```

<br>

## EDR

<br>

```{r categories by edr}
cat.edr <- original %>%
  filter(group %in% c(1,4)) %>%
  drop_na(edr) %>%
  group_by(edr) %>%
  summarize(bsrev = sum(bsrev),
            exdayr = sum(exdayr),
            gtrev = sum(gtrev), 
            smlrev = sum(smlrev), 
            decrev = sum(decrev), 
            locaid = sum(locaid), 
            afdc22 = sum(afdc22), 
            totlep = sum(totlep), 
            totspar = sum(totspar), 
            cexaid = sum(cexaid, na.rm = TRUE), 
            tsparrev = sum(tsparrev), 
            eqaid = sum(eqaid), 
            hhaid = sum(hhaid), 
            penrev = sum(penrev)) %>%
  gather(key = "categories", value = "revenue", 2:15) %>%
  group_by(edr) %>%
  mutate(pct.revenue = revenue / sum(revenue)) %>%
  ungroup() %>%
  mutate(data_id = as.character(seq(n()))) %>%
  left_join(var.codes, by = c("categories" = "Code")) %>%
  select(edr, Category, pct.revenue) %>%
  spread(edr, pct.revenue)

datatable(cat.edr, class = "cell-border stripe", filter = "top", rownames = FALSE,
          options = list(columnDefs = list(list(className = "dt-center", targets = 1:6)),
                         scrollX = TRUE)) %>%
  formatPercentage(2:14, digits = 1)


```

<br>

# Categories per APU{.tabset}

The table below provides each of the revenue streams per APU for each school district as well as the total revenue per APU. If sorting by which schools receive the highest total revenue per APU, the top ten is mostly composed of small rural school districts along with a few charter schools.

Of the categories that seem to drive the higher total revenue per APU are compensatory, sparsity, small schools, and transportation sparsity.

<br>

```{r categories per apu districts}
cat.total.rev.apu.districts <- original %>%
  filter(group %in% c(1,4),
         `District Name` != "Minnesota State Academies") %>%
  select(`District Name`, group, awadm22, all_of(categories)) %>%
  gather(key = "categories", value = "value", 3:17) %>%
  group_by(`District Name`, group) %>%
  summarize(awadm22 = value[categories == "awadm22"],
            revenue = sum(value[categories != "awadm22"])) %>%
  ungroup() %>%
  mutate(rev.awadm22 = revenue / awadm22)
  
cat.apu.districts <- original %>%
  filter(group %in% c(1,4),
         `District Name` != "Minnesota State Academies") %>%
  select(`District Name`, group, awadm22, all_of(categories)) %>%
  gather(key = "categories", value = "revenue", 4:17) %>%
  mutate(revenue = ifelse(is.na(revenue), 0, revenue),
         revenue.apu = round(revenue / awadm22)) %>%
  left_join(var.codes, by = c("categories" = "Code")) %>%
  left_join(group.codes, by = c("group" = "Code")) %>%
  select(`District Name`, Definition, Category, revenue.apu) %>%
  spread(Category, revenue.apu) %>%
  left_join(cat.total.rev.apu.districts, by = c("District Name")) %>%
  select(`District Name`, Definition, awadm22, revenue, rev.awadm22, 3:16) %>%
  rename(APU = awadm22,
         `Total revenue` = revenue,
         `Total revenue per APU` = rev.awadm22)

datatable(cat.apu.districts, class = "cell-border stripe", filter = "top", rownames = FALSE,
          options = list(columnDefs = list(list(className = "dt-center", targets = 1:18)),
                         scrollX = TRUE)) %>%
  formatCurrency(3, "", digits = 0) %>%
  formatCurrency(4:19, "$", digits = 0)
```

<br>

The most important part is to see if there are significant differences in the amount of revenue per APU based on the type of school, RUCA category, and regions. Let's take a look.

We will begin with the types of schools.

Out of the 14 categories, public schools received a higher revenue per APU ratio for 5 - extended time, local optional aid, pensions adjustment, sparsity, and transportation sparsity. Charter schools received a higher revenue per APU ratio for 7 categories.

Some of the categories with significant differences are compensatory, english learners, and small schools revenue.

<br>

```{r categories per apu group}
cat.total.rev.apu.group <- original %>%
  filter(group %in% c(1,4),
         `District Name` != "Minnesota State Academies") %>%
  select(group, awadm22, all_of(categories)) %>%
  gather(key = "categories", value = "value", 2:16) %>%
  group_by(group) %>%
  summarize(awadm22 = round(sum(value[categories == "awadm22"], na.rm = TRUE)),
            revenue = round(sum(value[categories != "awadm22"], na.rm = TRUE))) %>%
  ungroup() %>%
  mutate(rev.awadm22 = round(revenue / awadm22)) %>%
  left_join(group.codes, by = c("group" = "Code")) %>%
  select(Definition, awadm22, revenue, rev.awadm22)
  
cat.apu.group <- original %>%
  filter(group %in% c(1,4),
         `District Name` != "Minnesota State Academies") %>%
  select(group, awadm22, all_of(categories)) %>%
  gather(key = "categories", value = "revenue", 2:16) %>%
  group_by(group, categories) %>%
  summarize(value = sum(revenue, na.rm = TRUE)) %>%
  ungroup() %>%
  group_by(group) %>%
  mutate(rev.apu = round(value / value[categories=="awadm22"])) %>%
  ungroup() %>%
  filter(categories != "awadm22") %>%
  left_join(var.codes, by = c("categories" = "Code")) %>%
  left_join(group.codes, by = c("group" = "Code")) %>%
  select(Definition, Category, rev.apu) %>%
  spread(Category, rev.apu) %>%
  left_join(cat.total.rev.apu.group, by = c("Definition")) %>%
  select(Definition, awadm22, revenue, rev.awadm22, 2:15) %>%
  rename(APU = awadm22,
         `Total revenue` = revenue,
         `Total revenue per APU` = rev.awadm22)

datatable(cat.apu.group, class = "cell-border stripe", filter = "top", rownames = FALSE,
          options = list(columnDefs = list(list(className = "dt-center", targets = 1:17)),
                         scrollX = TRUE)) %>%
  formatStyle(columns = "Total revenue per APU", backgroundColor = styleEqual(max(cat.apu.group$`Total revenue per APU`), "#a6bddb")) %>%
  formatStyle(columns = "Compensatory", backgroundColor = styleEqual(max(cat.apu.group$Compensatory), "#a6bddb")) %>%
  formatStyle(columns = "English learners", backgroundColor = styleEqual(max(cat.apu.group$`English learners`), "#a6bddb")) %>%
  formatStyle(columns = "Declining enrollment", backgroundColor = styleEqual(max(cat.apu.group$`Declining enrollment`), "#a6bddb")) %>%
  formatStyle(columns = "Equity aid", backgroundColor = styleEqual(max(cat.apu.group$`Equity aid`), "#a6bddb")) %>%
  formatStyle(columns = "Extended time", backgroundColor = styleEqual(max(cat.apu.group$`Extended time`), "#a6bddb")) %>%
  formatStyle(columns = "Local optional aid", backgroundColor = styleEqual(max(cat.apu.group$`Local optional aid`), "#a6bddb")) %>%
  formatStyle(columns = "Operating capital aid", backgroundColor = styleEqual(max(cat.apu.group$`Operating capital aid`), "#a6bddb")) %>%
  formatStyle(columns = "Pensions adjustment", backgroundColor = styleEqual(max(cat.apu.group$`Pensions adjustment`), "#a6bddb")) %>%
  formatStyle(columns = "Small schools", backgroundColor = styleEqual(max(cat.apu.group$`Small schools`), "#a6bddb")) %>%
  formatStyle(columns = "Sparsity", backgroundColor = styleEqual(max(cat.apu.group$Sparsity), "#a6bddb")) %>%
  formatStyle(columns = "Transition aid", backgroundColor = styleEqual(max(cat.apu.group$`Transition aid`), "#a6bddb")) %>%
  formatStyle(columns = "Transportation sparsity", backgroundColor = styleEqual(max(cat.apu.group$`Transportation sparsity`), "#a6bddb")) %>%
  formatCurrency(2, "", digits = 0) %>%
  formatCurrency(3:18, "$", digits = 0)
```

<br>

Next, let's look at differences across RUCA categories and regions.

The RUCA table below shows that schools in entirely rural counties receive the most revenue per APU ($8,903) compared to the other RUCA categories. Entirely urban receive the least amount with $7,753.

The categories making the biggest difference for rural schools are equity aid, local optional aid, small schools, sparsity, and transportation sparsity. 

The planning regions table shows that our three most rural regions - Northeast, Northwest, and Southwest, have very similar total revenue per APU - all have just a little over $8,000 per APU. The lowest is in the seven county metro and Central regions. There higher revenue per APU is driven by the same categories as rural county groups - sparsity, equity, local optional and small schools.

The top 5 EDRs all represent their regions order as well - EDR 2, EDR 1, EDR 6W and EDR 3 are all in the Northwest, Northeast, and Southwest regions. EDR 2 has a high revenue per APU due to the crazy high compensatory revenue per APU. 

Interestingly, the lowest are comprised of somewhat rural areas - EDR 7W, EDR 7E, EDR 9, and EDR 4. These lower revenues per APU are driven by lower revenues from sparsity, compensatory, and equity aid.



<br>

## RUCA

<br>

```{r categories per apu ruca}
cat.total.rev.apu.ruca <- original %>%
  filter(group %in% c(1,4),
         `District Name` != "Minnesota State Academies") %>%
  select(Dem_Desc, awadm22, all_of(categories)) %>%
  gather(key = "categories", value = "value", 2:16) %>%
  group_by(Dem_Desc) %>%
  summarize(awadm22 = round(sum(value[categories == "awadm22"], na.rm = TRUE)),
            revenue = round(sum(value[categories != "awadm22"], na.rm = TRUE))) %>%
  ungroup() %>%
  mutate(rev.awadm22 = round(revenue / awadm22))
  
cat.apu.ruca <- original %>%
  filter(group %in% c(1,4),
         `District Name` != "Minnesota State Academies") %>%
  drop_na(Dem_Desc) %>%
  select(Dem_Desc, awadm22, all_of(categories)) %>%
  gather(key = "categories", value = "revenue", 2:16) %>%
  group_by(Dem_Desc, categories) %>%
  summarize(value = sum(revenue, na.rm = TRUE)) %>%
  ungroup() %>%
  group_by(Dem_Desc) %>%
  mutate(rev.apu = round(value / value[categories=="awadm22"])) %>%
  ungroup() %>%
  filter(categories != "awadm22") %>%
  left_join(var.codes, by = c("categories" = "Code")) %>%
  select(Dem_Desc, Category, rev.apu) %>%
  spread(Category, rev.apu) %>%
  left_join(cat.total.rev.apu.ruca, by = "Dem_Desc") %>%
  select(Dem_Desc, awadm22, revenue, rev.awadm22, 2:15) %>%
  rename(APU = awadm22,
         `Total revenue` = revenue,
         `Total revenue per APU` = rev.awadm22)

datatable(cat.apu.ruca, class = "cell-border stripe", filter = "top", rownames = FALSE,
          options = list(columnDefs = list(list(className = "dt-center", targets = 1:17)),
                         scrollX = TRUE)) %>%
  formatStyle(columns = "Total revenue per APU", backgroundColor = styleEqual(max(cat.apu.ruca$`Total revenue per APU`), "#a6bddb")) %>%
  formatStyle(columns = "Compensatory", backgroundColor = styleEqual(max(cat.apu.ruca$Compensatory), "#a6bddb")) %>%
  formatStyle(columns = "English learners", backgroundColor = styleEqual(max(cat.apu.ruca$`English learners`), "#a6bddb")) %>%
  formatStyle(columns = "Declining enrollment", backgroundColor = styleEqual(max(cat.apu.ruca$`Declining enrollment`), "#a6bddb")) %>%
  formatStyle(columns = "Equity aid", backgroundColor = styleEqual(max(cat.apu.ruca$`Equity aid`), "#a6bddb")) %>%
  formatStyle(columns = "Extended time", backgroundColor = styleEqual(max(cat.apu.ruca$`Extended time`), "#a6bddb")) %>%
  formatStyle(columns = "Local optional aid", backgroundColor = styleEqual(max(cat.apu.ruca$`Local optional aid`), "#a6bddb")) %>%
  formatStyle(columns = "Operating capital aid", backgroundColor = styleEqual(max(cat.apu.ruca$`Operating capital aid`), "#a6bddb")) %>%
  formatStyle(columns = "Pensions adjustment", backgroundColor = styleEqual(max(cat.apu.ruca$`Pensions adjustment`), "#a6bddb")) %>%
  formatStyle(columns = "Small schools", backgroundColor = styleEqual(max(cat.apu.ruca$`Small schools`), "#a6bddb")) %>%
  formatStyle(columns = "Sparsity", backgroundColor = styleEqual(max(cat.apu.ruca$Sparsity), "#a6bddb")) %>%
  formatStyle(columns = "Transition aid", backgroundColor = styleEqual(max(cat.apu.ruca$`Transition aid`), "#a6bddb")) %>%
  formatStyle(columns = "Transportation sparsity", backgroundColor = styleEqual(max(cat.apu.ruca$`Transportation sparsity`), "#a6bddb")) %>%
  formatCurrency(2, "", digits = 0) %>%
  formatCurrency(3:18, "$", digits = 0)
```

<br>

## Planning Region

<br>

```{r categories per apu pr}
cat.total.rev.apu.pr <- original %>%
  filter(group %in% c(1,4),
         `District Name` != "Minnesota State Academies") %>%
  select(planning.region, awadm22, all_of(categories)) %>%
  gather(key = "categories", value = "value", 2:16) %>%
  group_by(planning.region) %>%
  summarize(awadm22 = round(sum(value[categories == "awadm22"], na.rm = TRUE)),
            revenue = round(sum(value[categories != "awadm22"], na.rm = TRUE))) %>%
  ungroup() %>%
  mutate(rev.awadm22 = round(revenue / awadm22))
  
cat.apu.pr <- original %>%
  filter(group %in% c(1,4),
         `District Name` != "Minnesota State Academies") %>%
  drop_na(planning.region) %>%
  select(planning.region, awadm22, all_of(categories)) %>%
  gather(key = "categories", value = "revenue", 2:16) %>%
  group_by(planning.region, categories) %>%
  summarize(value = sum(revenue, na.rm = TRUE)) %>%
  ungroup() %>%
  group_by(planning.region) %>%
  mutate(rev.apu = round(value / value[categories=="awadm22"])) %>%
  ungroup() %>%
  filter(categories != "awadm22") %>%
  left_join(var.codes, by = c("categories" = "Code")) %>%
  select(planning.region, Category, rev.apu) %>%
  spread(Category, rev.apu) %>%
  left_join(cat.total.rev.apu.pr, by = "planning.region") %>%
  select(planning.region, awadm22, revenue, rev.awadm22, 2:15) %>%
  rename(APU = awadm22,
         `Total revenue` = revenue,
         `Total revenue per APU` = rev.awadm22)

datatable(cat.apu.pr, class = "cell-border stripe", filter = "top", rownames = FALSE,
          options = list(columnDefs = list(list(className = "dt-center", targets = 1:17)),
                         scrollX = TRUE)) %>%
  formatStyle(columns = "Total revenue per APU", backgroundColor = styleEqual(max(cat.apu.pr$`Total revenue per APU`), "#a6bddb")) %>%
  formatStyle(columns = "Compensatory", backgroundColor = styleEqual(max(cat.apu.pr$Compensatory), "#a6bddb")) %>%
  formatStyle(columns = "English learners", backgroundColor = styleEqual(max(cat.apu.pr$`English learners`), "#a6bddb")) %>%
  formatStyle(columns = "Declining enrollment", backgroundColor = styleEqual(max(cat.apu.pr$`Declining enrollment`), "#a6bddb")) %>%
  formatStyle(columns = "Equity aid", backgroundColor = styleEqual(max(cat.apu.pr$`Equity aid`), "#a6bddb")) %>%
  formatStyle(columns = "Extended time", backgroundColor = styleEqual(max(cat.apu.pr$`Extended time`), "#a6bddb")) %>%
  formatStyle(columns = "Local optional aid", backgroundColor = styleEqual(max(cat.apu.pr$`Local optional aid`), "#a6bddb")) %>%
  formatStyle(columns = "Operating capital aid", backgroundColor = styleEqual(max(cat.apu.pr$`Operating capital aid`), "#a6bddb")) %>%
  formatStyle(columns = "Pensions adjustment", backgroundColor = styleEqual(max(cat.apu.pr$`Pensions adjustment`), "#a6bddb")) %>%
  formatStyle(columns = "Small schools", backgroundColor = styleEqual(max(cat.apu.pr$`Small schools`), "#a6bddb")) %>%
  formatStyle(columns = "Sparsity", backgroundColor = styleEqual(max(cat.apu.pr$Sparsity), "#a6bddb")) %>%
  formatStyle(columns = "Transition aid", backgroundColor = styleEqual(max(cat.apu.pr$`Transition aid`), "#a6bddb")) %>%
  formatStyle(columns = "Transportation sparsity", backgroundColor = styleEqual(max(cat.apu.pr$`Transportation sparsity`), "#a6bddb")) %>%
  formatCurrency(2, "", digits = 0) %>%
  formatCurrency(3:18, "$", digits = 0)
```

<br>

## EDR

<br>

```{r categories per apu edr}
cat.total.rev.apu.edr <- original %>%
  filter(group %in% c(1,4),
         `District Name` != "Minnesota State Academies") %>%
  select(edr, awadm22, all_of(categories)) %>%
  gather(key = "categories", value = "value", 2:16) %>%
  group_by(edr) %>%
  summarize(awadm22 = round(sum(value[categories == "awadm22"], na.rm = TRUE)),
            revenue = round(sum(value[categories != "awadm22"], na.rm = TRUE))) %>%
  ungroup() %>%
  mutate(rev.awadm22 = round(revenue / awadm22))
  
cat.apu.edr <- original %>%
  filter(group %in% c(1,4),
         `District Name` != "Minnesota State Academies") %>%
  drop_na(edr) %>%
  select(edr, awadm22, all_of(categories)) %>%
  gather(key = "categories", value = "revenue", 2:16) %>%
  group_by(edr, categories) %>%
  summarize(value = sum(revenue, na.rm = TRUE)) %>%
  ungroup() %>%
  group_by(edr) %>%
  mutate(rev.apu = round(value / value[categories=="awadm22"])) %>%
  ungroup() %>%
  filter(categories != "awadm22") %>%
  left_join(var.codes, by = c("categories" = "Code")) %>%
  select(edr, Category, rev.apu) %>%
  spread(Category, rev.apu) %>%
  left_join(cat.total.rev.apu.edr, by = "edr") %>%
  select(edr, awadm22, revenue, rev.awadm22, 2:15) %>%
  rename(APU = awadm22,
         `Total revenue` = revenue,
         `Total revenue per APU` = rev.awadm22)

datatable(cat.apu.edr, class = "cell-border stripe", filter = "top", rownames = FALSE,
          options = list(columnDefs = list(list(className = "dt-center", targets = 1:17)),
                         scrollX = TRUE)) %>%
  formatStyle(columns = "Total revenue per APU", backgroundColor = styleEqual(max(cat.apu.edr$`Total revenue per APU`), "#a6bddb")) %>%
  formatStyle(columns = "Compensatory", backgroundColor = styleEqual(max(cat.apu.edr$Compensatory), "#a6bddb")) %>%
  formatStyle(columns = "English learners", backgroundColor = styleEqual(max(cat.apu.edr$`English learners`), "#a6bddb")) %>%
  formatStyle(columns = "Declining enrollment", backgroundColor = styleEqual(max(cat.apu.edr$`Declining enrollment`), "#a6bddb")) %>%
  formatStyle(columns = "Equity aid", backgroundColor = styleEqual(max(cat.apu.edr$`Equity aid`), "#a6bddb")) %>%
  formatStyle(columns = "Extended time", backgroundColor = styleEqual(max(cat.apu.edr$`Extended time`), "#a6bddb")) %>%
  formatStyle(columns = "Local optional aid", backgroundColor = styleEqual(max(cat.apu.edr$`Local optional aid`), "#a6bddb")) %>%
  formatStyle(columns = "Operating capital aid", backgroundColor = styleEqual(max(cat.apu.edr$`Operating capital aid`), "#a6bddb")) %>%
  formatStyle(columns = "Pensions adjustment", backgroundColor = styleEqual(max(cat.apu.edr$`Pensions adjustment`), "#a6bddb")) %>%
  formatStyle(columns = "Small schools", backgroundColor = styleEqual(max(cat.apu.edr$`Small schools`), "#a6bddb")) %>%
  formatStyle(columns = "Sparsity", backgroundColor = styleEqual(max(cat.apu.edr$Sparsity), "#a6bddb")) %>%
  formatStyle(columns = "Transition aid", backgroundColor = styleEqual(max(cat.apu.edr$`Transition aid`), "#a6bddb")) %>%
  formatStyle(columns = "Transportation sparsity", backgroundColor = styleEqual(max(cat.apu.edr$`Transportation sparsity`), "#a6bddb")) %>%
  formatCurrency(2, "", digits = 0) %>%
  formatCurrency(3:18, "$", digits = 0)
```

<br>




