---
title: "Basic education revenue"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 4
runtime: shiny
resource_files:
- Data/Shapefiles/County shapefiles/MNCounties_MNDOT.cpg
- Data/Shapefiles/County shapefiles/MNCounties_MNDOT.dbf
- Data/Shapefiles/County shapefiles/MNCounties_MNDOT.prj
- Data/Shapefiles/County shapefiles/MNCounties_MNDOT.sbn
- Data/Shapefiles/County shapefiles/MNCounties_MNDOT.sbx
- Data/Shapefiles/County shapefiles/MNCounties_MNDOT.shp.xml
- Data/Shapefiles/County shapefiles/MNCounties_MNDOT.shx
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(tidyverse)
library(sf)
library(ggrepel)
library(scales)
library(shiny)
library(shinycssloaders)
library(ggiraph)
library(kableExtra)
library(rmapshaper)
library(cowplot)
library(DT)
library(htmlwidgets)
library(RColorBrewer)
library(readxl)
library(janitor)
library(lubridate)
library(systemfonts)
reset_font_cache()
library(ggtext)

```

```{r loading jon docs and shapefiles, cache=TRUE, include=FALSE}
theme_bar <- theme_bw() +
  theme(panel.grid.major = element_line(color = "grey70", size = 0.1),
        panel.grid.minor = element_blank(),
        axis.ticks = element_blank(),
        axis.text = element_text(face = "bold"),
        panel.border = element_blank(),
        legend.background = element_rect(fill = "transparent", color = "transparent"),
        legend.key = element_rect(fill = "transparent"),
        legend.key.size = unit(1, "lines"),
        legend.margin = margin(0,0,0,0),
        legend.title = element_blank(),
        legend.text = element_text(margin = margin(l = 2)),
        text = element_text(family = "Arial") ,
        plot.title.position = "plot",
        plot.title = element_text(face = "bold"))

theme_line <- theme_bw() +
  theme(legend.background = element_rect(fill = "transparent", color = "transparent"),
        legend.key = element_rect(fill = "transparent"),
        legend.text = element_text(margin = margin(l = 2)),
        panel.grid.minor = element_blank(),
        panel.grid.major = element_line(color = "grey70", size = 0.1),
        axis.ticks = element_blank(),
        axis.text = element_text(face = "bold"),
        panel.border = element_blank(),
        legend.margin = margin(0,0,0,0),
        legend.key.size = unit(1, "lines"),
        text = element_text(family = "Arial") ,
        plot.title.position = "plot",
        plot.title = element_text(face = "bold"))


theme_sf <- theme_bw() +
  theme(axis.text.x=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks=element_blank(),
        panel.background = element_blank(),
        panel.grid.major = element_line(color = "white"),
        panel.border = element_blank(),
        legend.title = element_blank(),
        legend.text = element_text(margin = margin(l = 2)),
        legend.margin = margin(0,0,0,0),
        legend.key.size = unit(1, "lines"),
        text = element_text(family = "Arial") ,
        plot.title.position = "plot",
        plot.title = element_text(face = "bold"))

regions <- read_csv("Data/Join docs/county_regions.csv") %>%
    select(5,6) %>%
    unique() %>%
    mutate(edr = str_replace(edr, "  ", " "),
           planning.region = str_replace(planning.region, " Minnesota", ""),
           planning.region = fct_relevel(planning.region, "Northwest", "Northeast", "Central", "Seven County Mpls-St Paul", "Southwest", "Southeast"),
           edr = fct_relevel(edr, "EDR 1 - Northwest", "EDR 2 - Headwaters", "EDR 3 - Arrowhead", "EDR 4 - West Central", "EDR 5 - North Central", "EDR 6E- Southwest Central", "EDR 6W- Upper Minnesota Valley", "EDR 7E- East Central", "EDR 7W- Central", "EDR 8 - Southwest", "EDR 9 - South Central", "EDR 10 - Southeast", "EDR 11 - 7 County Twin Cities", "Minnesota"))

counties.regions <- read_csv("Data/Join docs/county_regions.csv") %>%
  rename(mif = `MIF Region`) %>%
  mutate(countyfp = formatC(countyfp, width = 3, flag = "0"),
         Name = str_to_title(Name),
         Name = str_replace(Name, "Q", "q"),
         Name = str_replace(Name, "Of The", "of the"),
         Name = str_replace(Name, "Mcleod", "McLeod"),
         Dem_Desc = ifelse(Name == "Minnesota", "Minnesota", Dem_Desc) ,
         edr = str_replace(edr, "  ", " "),
         planning.region = str_replace(planning.region, " Minnesota", ""),
         planning.region = fct_relevel(planning.region, "Northwest", "Northeast", "Central", "Seven County Mpls-St Paul", "Southwest", "Southeast"),
         edr = fct_relevel(edr, "EDR 1 - Northwest", "EDR 2 - Headwaters", "EDR 3 - Arrowhead", "EDR 4 - West Central", "EDR 5 - North Central", "EDR 6E- Southwest Central", "EDR 6W- Upper Minnesota Valley", "EDR 7E- East Central", "EDR 7W- Central", "EDR 8 - Southwest", "EDR 9 - South Central", "EDR 10 - Southeast", "EDR 11 - 7 County Twin Cities", "Minnesota"),
         mif = ifelse(is.na(mif), "TC", mif),
         mif = as.factor(mif),
         mif = fct_relevel(mif, "NW", "NE", "WC", "EC", "SW", "SE", "TC"))


color.ruca <- c("Entirely rural" = "#009933", "Town/rural mix" = "#99CC33", "Urban/town/rural mix" = "#CC9966", "Entirely urban" = "#754C29", "Minnesota" = "black")

color.pr <- c("Northwest" = 	"#4575b4", "Northeast" = "grey", "Central" = "#fee090", "Seven County Mpls-St Paul" = "#d73027", "Southwest" = "#91bfdb", "Southeast" = "#fc8d59", "Minnesota" = "black")

color.edr <- c("EDR 1 - Northwest" = "#b3cde3", "EDR 2 - Headwaters" = "#8c96c6", "EDR 3 - Arrowhead" = "#fe9929", "EDR 4 - West Central" = "#8856a7", "EDR 5 - North Central" = "#810f7c", "EDR 6E- Southwest Central" = "#e5f5f9", "EDR 6W- Upper Minnesota Valley" = "#bdc9e1", "EDR 7E- East Central" = "#99d8c9", "EDR 7W- Central" = "#2ca25f", "EDR 8 - Southwest" = "#74a9cf", "EDR 9 - South Central" = "#0570b0", "EDR 10 - Southeast" = "#d7301f", "EDR 11 - 7 County Twin Cities" = "#d8b365", "Minnesota" = "black")

color.pr.edr <- c ("Northwest" = "#4575b4","Northeast" = "#e0f3f8", "Central" = "#fee090", "Seven County Mpls-St Paul" = "#d73027", "Southwest" = "#91bfdb", "Southeast" = "#fc8d59", "Minnesota" = "black", "EDR 1 - Northwest" = "#b3cde3", "EDR 2 - Headwaters" = "#8c96c6", "EDR 3 - Arrowhead" = "#fe9929", "EDR 4 - West Central" = "#8856a7", "EDR 5 - North Central" = "#810f7c", "EDR 6E- Southwest Central" = "#e5f5f9", "EDR 6W- Upper Minnesota Valley" = "#bdc9e1", "EDR 7E- East Central" = "#99d8c9", "EDR 7W- Central" = "#2ca25f", "EDR 8 - Southwest" = "#74a9cf", "EDR 9 - South Central" = "#0570b0", "EDR 10 - Southeast" = "#d7301f", "EDR 11 - 7 County Twin Cities" = "#d8b365")

mn_counties <- st_read("Data/Shapefiles/county shapefiles/MNCounties_MNDOT.shp", quiet = TRUE) %>%
  ms_simplify(keep = .01, keep_shapes = TRUE) %>%
  rename(countyfp = FIPS_CODE)

```


```{r master basic revenue}
original <- read_csv("Data/Formula/Master.csv") %>%
  mutate(Dem_Desc = fct_relevel(Dem_Desc, "Entirely rural", "Town/rural mix", "Urban/town/rural mix", "Entirely urban"),
         planning.region = fct_relevel(planning.region, "Northwest", "Northeast", "Central", "Seven County Mpls-St Paul", "Southwest", "Southeast"),
         edr = fct_relevel(edr, "EDR 1 - Northwest", "EDR 2 - Headwaters", "EDR 3 - Arrowhead", "EDR 4 - West Central", "EDR 5 - North Central", "EDR 6E- Southwest Central", "EDR 6W- Upper Minnesota Valley", "EDR 7E- East Central", "EDR 7W- Central", "EDR 8 - Southwest", "EDR 9 - South Central", "EDR 10 - Southeast", "EDR 11 - 7 County Twin Cities", "Minnesota"))

group.codes <- read_xlsx("Data/Formula/group codes.xlsx")

```

# Description

The basic education revenue category provides a base amount of revenue per adjusted pupil units to each school district. 

* FY23: $6,863
* FY22: $6,728
* FY21: $6,567
* FY20: $6,438
* FY19: $6,312
* FY18: $6,188

<br>

# State Legislation
[126C.10 Subd. 2.Basic revenue.](https://www.revisor.mn.gov/statutes/cite/126C.10)
 
The basic revenue for each district equals the formula allowance times the adjusted pupil units for the school year. The formula allowance for fiscal year 2021 is $6,567. The formula allowance for fiscal year 2022 is $6,728. The formula allowance for fiscal year 2023 and later is $6,863.

<br>

# Data {.tabset}
Since the only official student enrollment data we have is for last year's school year, we will stick with the FY22 data as well as basic revenue value. 

The total basic revenue distributed by the state of Minnesota in 2021-2022 school year was `r dollar(sum(original$bsrev))`. 

The table below procides the basic revenue received by each school district in Minnesota.

<br>

```{r basic revenue table}
basic.rev.schools <- original %>%
  filter(group %in% c(1,4)) %>%
  select(`District Name`, dist, type, `County Name`, bsrev, awadm22) %>%
  rename(`District Number` = 2,
         `District Type` = 3,
         `Basic Revenue` = 5,
         `Adjusted Pupil Units` = 6) %>%
  mutate(`Basic revenue per pupil unit` = `Basic Revenue` / `Adjusted Pupil Units`)
  
datatable(basic.rev.schools, class = "cell-border stripe", filter = "top", rownames = FALSE,
          options = list(columnDefs = list(list(className = "dt-center", targets = 1:4)))) %>%
  formatCurrency(5, "$", digits = 2) %>%
  formatCurrency(6, "", digits = 0) %>%
  formatCurrency(7, "$", digits = 2)

```
<br>

Below provides a breakdown of the different types of education institutions that receive funding from the state and how much each of those types received in FY22. 

As the table shows, the basic revenue category provides a flat $6,728 for each adjusted pupil unit.

<br>

```{r basic revenue by school type}
basic.rev.school.type <- original %>%
  filter(group %in% c(1,4)) %>%
  group_by(group) %>%
  summarize(`Adjusted pupil units` = sum(awadm22, na.rm = TRUE),
            Districts = n(),
            `Basic Revenue` = sum(bsrev, na.rm = TRUE)) %>%
  ungroup() %>%
  mutate(`Basic revenue per district` = `Basic Revenue` / Districts,
         `Basic revenue per adjusted pupil` = `Basic Revenue` / `Adjusted pupil units`) %>%
  left_join(group.codes, by = c("group" = "Code")) %>%
  select(7, 2:6)

datatable(basic.rev.school.type, class = "cell-border stripe", filter = "top", rownames = FALSE,
          options = list(columnDefs = list(list(className = "dt-center", targets = 1:5)))) %>%
  formatCurrency(2:3, "", digits = 0) %>%
  formatCurrency(4:6, "$", digits = 0)
```

<br>

We should expect the exact same thing if we break it down by RUCA categories and regions.

<br>

## RUCA

<br>

```{r basic revenue ruca}
basic.rev.ruca <- original %>%
  filter(group %in% c(1,4)) %>%
  drop_na(Dem_Desc) %>%
  group_by(Dem_Desc) %>%
  summarize(`Adjusted pupil units` = sum(awadm22, na.rm = TRUE),
            Districts = n(),
            `Basic Revenue` = sum(bsrev, na.rm = TRUE)) %>%
  ungroup() %>%
  mutate(`Basic revenue per district` = `Basic Revenue` / Districts,
         `Basic revenue per adjusted pupil` = `Basic Revenue` / `Adjusted pupil units`) 

datatable(basic.rev.ruca, class = "cell-border stripe", filter = "top", rownames = FALSE,
          options = list(columnDefs = list(list(className = "dt-center", targets = 1:5)))) %>%
  formatCurrency(2:3, "", digits = 0) %>%
  formatCurrency(4:6, "$", digits = 0)

```

<br>

## Planning Regoion

<br>

```{r basic revenue pr}
basic.rev.pr <- original %>%
  filter(group %in% c(1,4)) %>%
  drop_na(planning.region) %>%
  group_by(planning.region) %>%
  summarize(`Adjusted pupil units` = sum(awadm22, na.rm = TRUE),
            Districts = n(),
            `Basic Revenue` = sum(bsrev, na.rm = TRUE)) %>%
  ungroup() %>%
  mutate(`Basic revenue per district` = `Basic Revenue` / Districts,
         `Basic revenue per adjusted pupil` = `Basic Revenue` / `Adjusted pupil units`) 

datatable(basic.rev.pr, class = "cell-border stripe", filter = "top", rownames = FALSE,
          options = list(columnDefs = list(list(className = "dt-center", targets = 1:5)))) %>%
  formatCurrency(2:3, "", digits = 0) %>%
  formatCurrency(4:6, "$", digits = 0)

```

<br>

## EDR

<br>

```{r basic revenue edr}
basic.rev.edr <- original %>%
  filter(group %in% c(1,4)) %>%
  drop_na(edr) %>%
  group_by(edr) %>%
  summarize(`Adjusted pupil units` = sum(awadm22, na.rm = TRUE),
            Districts = n(),
            `Basic Revenue` = sum(bsrev, na.rm = TRUE)) %>%
  ungroup() %>%
  mutate(`Basic revenue per district` = `Basic Revenue` / Districts,
         `Basic revenue per adjusted pupil` = `Basic Revenue` / `Adjusted pupil units`) 

datatable(basic.rev.edr, class = "cell-border stripe", filter = "top", rownames = FALSE,
          options = list(columnDefs = list(list(className = "dt-center", targets = 1:5)))) %>%
  formatCurrency(2:3, "", digits = 0) %>%
  formatCurrency(4:6, "$", digits = 0)

```


