---
title: "Equity"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 4
runtime: shiny
resource_files:
- Data/Shapefiles/County shapefiles/MNCounties_MNDOT.cpg
- Data/Shapefiles/County shapefiles/MNCounties_MNDOT.dbf
- Data/Shapefiles/County shapefiles/MNCounties_MNDOT.prj
- Data/Shapefiles/County shapefiles/MNCounties_MNDOT.sbn
- Data/Shapefiles/County shapefiles/MNCounties_MNDOT.sbx
- Data/Shapefiles/County shapefiles/MNCounties_MNDOT.shp.xml
- Data/Shapefiles/County shapefiles/MNCounties_MNDOT.shx
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(tidyverse)
library(sf)
library(ggrepel)
library(scales)
library(shiny)
library(shinycssloaders)
library(ggiraph)
library(kableExtra)
library(rmapshaper)
library(cowplot)
library(DT)
library(htmlwidgets)
library(RColorBrewer)
library(readxl)
library(janitor)
library(lubridate)
library(systemfonts)
reset_font_cache()
library(ggtext)
library(gmodels)

```

```{r loading jon docs and shapefiles, cache=TRUE, include=FALSE}
theme_bar <- theme_bw() +
  theme(panel.grid.major = element_line(color = "grey70", size = 0.1),
        panel.grid.minor = element_blank(),
        axis.ticks = element_blank(),
        axis.text = element_text(face = "bold"),
        panel.border = element_blank(),
        legend.background = element_rect(fill = "transparent", color = "transparent"),
        legend.key = element_rect(fill = "transparent"),
        legend.key.size = unit(1, "lines"),
        legend.margin = margin(0,0,0,0),
        legend.title = element_blank(),
        legend.text = element_text(margin = margin(l = 2)),
        text = element_text(family = "Arial"))

theme_line <- theme_bw() +
  theme(legend.background = element_rect(fill = "transparent", color = "transparent"),
        legend.key = element_rect(fill = "transparent"),
        legend.text = element_text(margin = margin(l = 2)),
        panel.grid.minor = element_blank(),
        panel.grid.major = element_line(color = "grey70", size = 0.1),
        axis.ticks = element_blank(),
        axis.text = element_text(face = "bold"),
        panel.border = element_blank(),
        legend.margin = margin(0,0,0,0),
        legend.key.size = unit(1, "lines"),
        text = element_text(family = "Arial"))


theme_sf <- theme_bw() +
  theme(axis.text.x=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks=element_blank(),
        panel.background = element_blank(),
        panel.grid.major = element_line(color = "white"),
        panel.border = element_blank(),
        plot.title = element_text(hjust = 0.5),
        legend.title = element_blank(),
        legend.text = element_text(margin = margin(l = 2)),
        legend.margin = margin(0,0,0,0),
        legend.key.size = unit(1, "lines"),
        text = element_text(family = "Arial"))

regions <- read_csv("Data/Join docs/county_regions.csv") %>%
    select(5,6) %>%
    unique() %>%
    mutate(edr = str_replace(edr, "  ", " "),
           planning.region = str_replace(planning.region, " Minnesota", ""),
           planning.region = fct_relevel(planning.region, "Northwest", "Northeast", "Central", "Seven County Mpls-St Paul", "Southwest", "Southeast"),
           edr = fct_relevel(edr, "EDR 1 - Northwest", "EDR 2 - Headwaters", "EDR 3 - Arrowhead", "EDR 4 - West Central", "EDR 5 - North Central", "EDR 6E- Southwest Central", "EDR 6W- Upper Minnesota Valley", "EDR 7E- East Central", "EDR 7W- Central", "EDR 8 - Southwest", "EDR 9 - South Central", "EDR 10 - Southeast", "EDR 11 - 7 County Twin Cities", "Minnesota"))

counties.regions <- read_csv("Data/Join docs/county_regions.csv") %>%
  rename(mif = `MIF Region`) %>%
  mutate(countyfp = formatC(countyfp, width = 3, flag = "0"),
         Name = str_to_title(Name),
         Name = str_replace(Name, "Q", "q"),
         Name = str_replace(Name, "Of The", "of the"),
         Name = str_replace(Name, "Mcleod", "McLeod"),
         planning.region = ifelse(Name == "Minnesota", "Minnesota", planning.region) ,
         edr = str_replace(edr, "  ", " "),
         planning.region = str_replace(planning.region, " Minnesota", ""),
         planning.region = fct_relevel(planning.region, "Northwest", "Northeast", "Central", "Seven County Mpls-St Paul", "Southwest", "Southeast"),
         edr = fct_relevel(edr, "EDR 1 - Northwest", "EDR 2 - Headwaters", "EDR 3 - Arrowhead", "EDR 4 - West Central", "EDR 5 - North Central", "EDR 6E- Southwest Central", "EDR 6W- Upper Minnesota Valley", "EDR 7E- East Central", "EDR 7W- Central", "EDR 8 - Southwest", "EDR 9 - South Central", "EDR 10 - Southeast", "EDR 11 - 7 County Twin Cities", "Minnesota"),
         mif = ifelse(is.na(mif), "TC", mif),
         mif = as.factor(mif),
         mif = fct_relevel(mif, "NW", "NE", "WC", "EC", "SW", "SE", "TC"))


color.edr <- c("Entirely rural" = "#009933", "Town/rural mix" = "#99CC33", "Urban/town/rural mix" = "#CC9966", "Entirely urban" = "#754C29", "Minnesota" = "black")

color.pr <- c("Northwest" = 	"#4575b4", "Northeast" = "grey", "Central" = "#fee090", "Seven County Mpls-St Paul" = "#d73027", "Southwest" = "#91bfdb", "Southeast" = "#fc8d59", "Minnesota" = "black")

color.edr <- c("EDR 1 - Northwest" = "#b3cde3", "EDR 2 - Headwaters" = "#8c96c6", "EDR 3 - Arrowhead" = "#fe9929", "EDR 4 - West Central" = "#8856a7", "EDR 5 - North Central" = "#810f7c", "EDR 6E- Southwest Central" = "#e5f5f9", "EDR 6W- Upper Minnesota Valley" = "#bdc9e1", "EDR 7E- East Central" = "#99d8c9", "EDR 7W- Central" = "#2ca25f", "EDR 8 - Southwest" = "#74a9cf", "EDR 9 - South Central" = "#0570b0", "EDR 10 - Southeast" = "#d7301f", "EDR 11 - 7 County Twin Cities" = "#d8b365", "Minnesota" = "black")

color.pr.edr <- c ("Northwest" = "#4575b4","Northeast" = "#e0f3f8", "Central" = "#fee090", "Seven County Mpls-St Paul" = "#d73027", "Southwest" = "#91bfdb", "Southeast" = "#fc8d59", "Minnesota" = "black", "EDR 1 - Northwest" = "#b3cde3", "EDR 2 - Headwaters" = "#8c96c6", "EDR 3 - Arrowhead" = "#fe9929", "EDR 4 - West Central" = "#8856a7", "EDR 5 - North Central" = "#810f7c", "EDR 6E- Southwest Central" = "#e5f5f9", "EDR 6W- Upper Minnesota Valley" = "#bdc9e1", "EDR 7E- East Central" = "#99d8c9", "EDR 7W- Central" = "#2ca25f", "EDR 8 - Southwest" = "#74a9cf", "EDR 9 - South Central" = "#0570b0", "EDR 10 - Southeast" = "#d7301f", "EDR 11 - 7 County Twin Cities" = "#d8b365")

mn_counties <- st_read("Data/Shapefiles/county shapefiles/MNCounties_MNDOT.shp", quiet = TRUE) %>%
  ms_simplify(keep = .01, keep_shapes = TRUE) %>%
  rename(countyfp = FIPS_CODE)

```

```{r master basic revenue}
original <- read_csv("Data/Formula/Master.csv") %>%
  mutate(Dem_Desc = fct_relevel(Dem_Desc, "Entirely rural", "Town/rural mix", "Urban/town/rural mix", "Entirely urban"),
         planning.region = fct_relevel(planning.region, "Northwest", "Northeast", "Central", "Seven County Mpls-St Paul", "Southwest", "Southeast"),
         edr = fct_relevel(edr, "EDR 1 - Northwest", "EDR 2 - Headwaters", "EDR 3 - Arrowhead", "EDR 4 - West Central", "EDR 5 - North Central", "EDR 6E- Southwest Central", "EDR 6W- Upper Minnesota Valley", "EDR 7E- East Central", "EDR 7W- Central", "EDR 8 - Southwest", "EDR 9 - South Central", "EDR 10 - Southeast", "EDR 11 - 7 County Twin Cities", "Minnesota"))

group.codes <- read_xlsx("Data/Formula/group codes.xlsx")

```

<br>

# Description

Equity revenue is designed to provide additional revenue to districts with lower amounts of referendum revenue. Calculations for this revenue is broken into two regions - the 7-county metro and greater Minnesota. The formula consists of three parts.

1. Basic equity revenue: equals the product of 125% x  [$14 +($80 x district’s equity index)] x adjusted pupil units
    + Equity index = 1 - (basic formula allowance + local optional revenue + referendum revenue) / per pupil amount for the district at the 95th percentile in that region
    + School districts of the first class (Minneapolis, St. Paul, and Duluth) don’t receive basic equity revenue
2. Low referendum revenue: district has per pupil referendum revenue less than 10% of the statewide average receives an additional equity amount equal to the lesser of $100,000 or the difference between 10% of the statewide average referendum revenue and the district’s current amount of referendum revenue.
3. Supplemental equity revenue: all school districts receive $50 per pupil unit.

Equity aid and levy: A district’s total equity revenue is equalized on referendum market value using an equalizing factor of $510,000.

<br>

# State Legislation
[Subd. 24.Equity revenue.](https://www.revisor.mn.gov/statutes/cite/126C.10#stat.126C.10.24)
 
a) A school district qualifies for equity revenue if:
    + the school district's adjusted pupil unit amount of basic revenue, transition revenue, first tier local optional revenue, and referendum revenue is less than the value of the school district at or immediately above the 95th percentile of school districts in its equity region for those revenue categories; and
    + the school district's administrative offices are not located in a city of the first class on July 1, 1999.
b) Equity revenue equals the product of
    + the district's adjusted pupil units for that year; times
    + the sum of (i) $14, plus (ii) $80, times the school district's equity index computed under subdivision 27.
c) A school district's equity revenue is increased by the greater of zero or an amount equal to the district's adjusted pupil units times the difference between ten percent of the statewide average amount of referendum revenue and first tier local optional revenue per adjusted pupil unit for that year and the sum of the district's referendum revenue and first tier local optional revenue per adjusted pupil unit. A school district's revenue under this paragraph must not exceed $100,000 for that year.
d) A school district's equity revenue for a school district located in the metro equity region equals the amount computed in paragraphs (b) and (c) multiplied by 1.25.
e) For fiscal year 2020 and later for a school district not included in paragraph (d), a district's equity revenue equals the amount computed in paragraphs (b) and (c) multiplied by 1.25.
f) A school district's additional equity revenue equals $50 times its adjusted pupil units.

[Subd. 25.Regional equity gap.](https://www.revisor.mn.gov/statutes/cite/126C.10#stat.126C.10.25)

The regional equity gap equals the difference between the value of the school district at or immediately above the fifth percentile of adjusted general revenue per adjusted pupil unit and the value of the school district at or immediately above the 95th percentile of adjusted general revenue per adjusted pupil unit.

[Subd. 26.District equity gap.](https://www.revisor.mn.gov/statutes/cite/126C.10#stat.126C.10.26)

A district's equity gap equals the greater of zero or the difference between the district's adjusted general revenue and the value of the school district at or immediately above the regional 95th percentile of adjusted general revenue per adjusted pupil unit.

[Subd. 27.District equity index.](https://www.revisor.mn.gov/statutes/cite/126C.10#stat.126C.10.27) 

A district's equity index equals the ratio of the sum of the district equity gap amount to the regional equity gap amount.

[Subd. 28.Equity region.](https://www.revisor.mn.gov/statutes/cite/126C.10#stat.126C.10.28)

For the purposes of computing equity revenue under subdivision 24, a district with its administrative office located in Anoka, Carver, Dakota, Hennepin, Ramsey, Scott, or Washington County on January 1, 2012, is part of the metro equity region. All other districts are part of the rural equity region.

[Subd. 29.Equity levy.](https://www.revisor.mn.gov/statutes/cite/126C.10#stat.126C.10.29)

To obtain equity revenue for fiscal year 2015 and later, a district may levy an amount not more than the product of its equity revenue for the fiscal year times the lesser of one or the ratio of its referendum market value per resident pupil unit to $510,000.

[Subd. 30.Equity aid.](https://www.revisor.mn.gov/statutes/cite/126C.10#stat.126C.10.30)

A district's equity aid equals its equity revenue minus its equity levy times the ratio of the actual amount levied to the permitted levy.

<br>

# Data

There are three parts to this component;

1. Equity revenue = eqrev
2. Equity levy = eqlvy
3. Equity aid = eqaid

Equity aid = equity revenue - equity levy

**Equity revenue**

*eqrev*

It's important to know that when comparing with schools in the 95th percentile, it's the comparison by region. There are two regions - seven county metro, and outside of seven county metro.

First, we need to figure out whether a school qualifies for equity funding. The calculations for this eligibility is laid out in Subd. 24a.

The formula below calculates the first part which is the sum of basic revenue, referendum revenue, transition revenue, and local optional revenue tier 1.

ajgened = bsrev + refrev22 + hhrev + loctier1rev

* bsrev = basic revenue
* refrev22 = referendum revenue
* hhrev = transition revenue
* loctier1rev = local optional revenue - first tier

It's then calculated at a per APU.

ajgeppu = ifelse(awadm22 > 0 & group == 1, round((ajgened / awadm22) * 100) / 100, ajgeppu)

Then, calculations are made to figure out the 95th and 5th percentile in each region. These calculations are then compared to ajgeppu and this get's the equity index.


 
 

<br>

```{r equity revenue district}
equity.districts <- original %>%
  filter(group %in% c(1,4)) %>%
  select(`District Name`, group, eqrev)

```