---
title: "Extended Time"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 4
runtime: shiny
resource_files:
- Data/Shapefiles/County shapefiles/MNCounties_MNDOT.cpg
- Data/Shapefiles/County shapefiles/MNCounties_MNDOT.dbf
- Data/Shapefiles/County shapefiles/MNCounties_MNDOT.prj
- Data/Shapefiles/County shapefiles/MNCounties_MNDOT.sbn
- Data/Shapefiles/County shapefiles/MNCounties_MNDOT.sbx
- Data/Shapefiles/County shapefiles/MNCounties_MNDOT.shp.xml
- Data/Shapefiles/County shapefiles/MNCounties_MNDOT.shx
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(tidyverse)
library(sf)
library(ggrepel)
library(scales)
library(shiny)
library(shinycssloaders)
library(ggiraph)
library(kableExtra)
library(rmapshaper)
library(cowplot)
library(DT)
library(htmlwidgets)
library(RColorBrewer)
library(readxl)
library(janitor)
library(lubridate)
library(systemfonts)
reset_font_cache()
library(ggtext)
library(gmodels)

```

```{r loading jon docs and shapefiles, cache=TRUE, include=FALSE}
theme_bar <- theme_bw() +
  theme(panel.grid.major = element_line(color = "grey70", size = 0.1),
        panel.grid.minor = element_blank(),
        axis.ticks = element_blank(),
        axis.text = element_text(face = "bold"),
        panel.border = element_blank(),
        legend.background = element_rect(fill = "transparent", color = "transparent"),
        legend.key = element_rect(fill = "transparent"),
        legend.key.size = unit(1, "lines"),
        legend.margin = margin(0,0,0,0),
        legend.title = element_blank(),
        legend.text = element_text(margin = margin(l = 2)),
        text = element_text(family = "Arial"))

theme_line <- theme_bw() +
  theme(legend.background = element_rect(fill = "transparent", color = "transparent"),
        legend.key = element_rect(fill = "transparent"),
        legend.text = element_text(margin = margin(l = 2)),
        panel.grid.minor = element_blank(),
        panel.grid.major = element_line(color = "grey70", size = 0.1),
        axis.ticks = element_blank(),
        axis.text = element_text(face = "bold"),
        panel.border = element_blank(),
        legend.margin = margin(0,0,0,0),
        legend.key.size = unit(1, "lines"),
        text = element_text(family = "Arial"))


theme_sf <- theme_bw() +
  theme(axis.text.x=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks=element_blank(),
        panel.background = element_blank(),
        panel.grid.major = element_line(color = "white"),
        panel.border = element_blank(),
        plot.title = element_text(hjust = 0.5),
        legend.title = element_blank(),
        legend.text = element_text(margin = margin(l = 2)),
        legend.margin = margin(0,0,0,0),
        legend.key.size = unit(1, "lines"),
        text = element_text(family = "Arial"))

regions <- read_csv("Data/Join docs/county_regions.csv") %>%
    select(5,6) %>%
    unique() %>%
    mutate(edr = str_replace(edr, "  ", " "),
           planning.region = str_replace(planning.region, " Minnesota", ""),
           planning.region = fct_relevel(planning.region, "Northwest", "Northeast", "Central", "Seven County Mpls-St Paul", "Southwest", "Southeast"),
           edr = fct_relevel(edr, "EDR 1 - Northwest", "EDR 2 - Headwaters", "EDR 3 - Arrowhead", "EDR 4 - West Central", "EDR 5 - North Central", "EDR 6E- Southwest Central", "EDR 6W- Upper Minnesota Valley", "EDR 7E- East Central", "EDR 7W- Central", "EDR 8 - Southwest", "EDR 9 - South Central", "EDR 10 - Southeast", "EDR 11 - 7 County Twin Cities", "Minnesota"))

counties.regions <- read_csv("Data/Join docs/county_regions.csv") %>%
  rename(mif = `MIF Region`) %>%
  mutate(countyfp = formatC(countyfp, width = 3, flag = "0"),
         Name = str_to_title(Name),
         Name = str_replace(Name, "Q", "q"),
         Name = str_replace(Name, "Of The", "of the"),
         Name = str_replace(Name, "Mcleod", "McLeod"),
         planning.region = ifelse(Name == "Minnesota", "Minnesota", planning.region) ,
         edr = str_replace(edr, "  ", " "),
         planning.region = str_replace(planning.region, " Minnesota", ""),
         planning.region = fct_relevel(planning.region, "Northwest", "Northeast", "Central", "Seven County Mpls-St Paul", "Southwest", "Southeast"),
         edr = fct_relevel(edr, "EDR 1 - Northwest", "EDR 2 - Headwaters", "EDR 3 - Arrowhead", "EDR 4 - West Central", "EDR 5 - North Central", "EDR 6E- Southwest Central", "EDR 6W- Upper Minnesota Valley", "EDR 7E- East Central", "EDR 7W- Central", "EDR 8 - Southwest", "EDR 9 - South Central", "EDR 10 - Southeast", "EDR 11 - 7 County Twin Cities", "Minnesota"),
         mif = ifelse(is.na(mif), "TC", mif),
         mif = as.factor(mif),
         mif = fct_relevel(mif, "NW", "NE", "WC", "EC", "SW", "SE", "TC"))


color.ruca <- c("Entirely rural" = "#009933", "Town/rural mix" = "#99CC33", "Urban/town/rural mix" = "#CC9966", "Entirely urban" = "#754C29", "Minnesota" = "black")

color.pr <- c("Northwest" = 	"#4575b4", "Northeast" = "grey", "Central" = "#fee090", "Seven County Mpls-St Paul" = "#d73027", "Southwest" = "#91bfdb", "Southeast" = "#fc8d59", "Minnesota" = "black")

color.edr <- c("EDR 1 - Northwest" = "#b3cde3", "EDR 2 - Headwaters" = "#8c96c6", "EDR 3 - Arrowhead" = "#fe9929", "EDR 4 - West Central" = "#8856a7", "EDR 5 - North Central" = "#810f7c", "EDR 6E- Southwest Central" = "#e5f5f9", "EDR 6W- Upper Minnesota Valley" = "#bdc9e1", "EDR 7E- East Central" = "#99d8c9", "EDR 7W- Central" = "#2ca25f", "EDR 8 - Southwest" = "#74a9cf", "EDR 9 - South Central" = "#0570b0", "EDR 10 - Southeast" = "#d7301f", "EDR 11 - 7 County Twin Cities" = "#d8b365", "Minnesota" = "black")

color.pr.edr <- c ("Northwest" = "#4575b4","Northeast" = "#e0f3f8", "Central" = "#fee090", "Seven County Mpls-St Paul" = "#d73027", "Southwest" = "#91bfdb", "Southeast" = "#fc8d59", "Minnesota" = "black", "EDR 1 - Northwest" = "#b3cde3", "EDR 2 - Headwaters" = "#8c96c6", "EDR 3 - Arrowhead" = "#fe9929", "EDR 4 - West Central" = "#8856a7", "EDR 5 - North Central" = "#810f7c", "EDR 6E- Southwest Central" = "#e5f5f9", "EDR 6W- Upper Minnesota Valley" = "#bdc9e1", "EDR 7E- East Central" = "#99d8c9", "EDR 7W- Central" = "#2ca25f", "EDR 8 - Southwest" = "#74a9cf", "EDR 9 - South Central" = "#0570b0", "EDR 10 - Southeast" = "#d7301f", "EDR 11 - 7 County Twin Cities" = "#d8b365")

mn_counties <- st_read("Data/Shapefiles/county shapefiles/MNCounties_MNDOT.shp", quiet = TRUE) %>%
  ms_simplify(keep = .01, keep_shapes = TRUE) %>%
  rename(countyfp = FIPS_CODE)

```

```{r master basic revenue}
original <- read_csv("Data/Formula/Master.csv") %>%
  mutate(Dem_Desc = fct_relevel(Dem_Desc, "Entirely rural", "Town/rural mix", "Urban/town/rural mix", "Entirely urban"),
         planning.region = fct_relevel(planning.region, "Northwest", "Northeast", "Central", "Seven County Mpls-St Paul", "Southwest", "Southeast"),
         edr = fct_relevel(edr, "EDR 1 - Northwest", "EDR 2 - Headwaters", "EDR 3 - Arrowhead", "EDR 4 - West Central", "EDR 5 - North Central", "EDR 6E- Southwest Central", "EDR 6W- Upper Minnesota Valley", "EDR 7E- East Central", "EDR 7W- Central", "EDR 8 - Southwest", "EDR 9 - South Central", "EDR 10 - Southeast", "EDR 11 - 7 County Twin Cities", "Minnesota"))

group.codes <- read_xlsx("Data/Formula/group codes.xlsx")

```

<br>

# Description

Beginning in 2004, school districts are prohibited from counting a student as more than 1.0 in average daily membership, which they used to do if a student was participating in a learning year program (summer school, extended day, extended week, etc..). This program allows a school district to count a student who participates in extended programming for up to an additional 0.2 students in ADM for the time the student spends in summer school, etc…. The allowance is $5,117 X the district’s extended time adjusted pupil units.

# State Legislation
[126C.10 Subd. 2.Basic revenue.](https://www.revisor.mn.gov/statutes/cite/126C.10#stat.126C.10.2a)

Subd. 2a.Extended time revenue.
 
a) A school district's extended time revenue is equal to the product of $5,117 and the sum of the adjusted pupil units of the district for each pupil in average daily membership in excess of 1.0 and less than 1.2 according to section [126C.05, subdivision 8](https://www.revisor.mn.gov/statutes/cite/126C.05#stat.126C.05.8).
	
b) Extended time revenue for pupils placed in an on-site education program at the Prairie Lakes Education Center or the Lake Park School, located within the borders of Independent School District No. 347, Willmar, for instruction provided after the end of the preceding regular school year and before the beginning of the following regular school year equals membership hours divided by the minimum annual instructional hours in section [126C.05, subdivision 15](https://www.revisor.mn.gov/statutes/cite/126C.05#stat.126C.05.15), not to exceed 0.20, times the pupil unit weighting in section [126C.05, subdivision 1](https://www.revisor.mn.gov/statutes/cite/126C.05#stat.126C.05.1), times $5,117.

c) A school district's extended time revenue may be used for extended day programs, extended week programs, summer school, vacation break academies such as spring break academies and summer term academies, and other programming authorized under the learning year program.

# Data

The following breakdowns use data from FY22. 

Extended time is a horizontal funding mechanism. Therefore it is distributed on a per student basis. In this case, it's based on students that participate in programs outside of the typical school day. Therefore, each school district will receive varying amounts depending on how many students they have participating in these programs, but on a per student basis, the revenue should equal $5,117 for each school district, RUCA category, and Region.

Below is the distribution of total extended time revenue provided to districts. A large majority of districts receive $0 in extended time revenue, indicating that they either don't provide any programs that qualify for extended time revenue or don't have any students participating in programs that qualify. 

Outside of this, most schools receive $500,000 or less.

<br>

```{r extended time pupil units distribution}
ext.time.distribution <- original %>%
  filter(group %in% c(1,4)) %>%
  select(dist, exdayr, awadm22) %>%
  mutate(ext.time.apu = exdayr / awadm22)

ext.time.distribution.plot <- ggplot(ext.time.distribution, aes(exdayr)) +
  geom_histogram() +
  labs(x="", y = "", color="", title = "Extended time revenue per APU - distribution")+
  scale_y_continuous(labels=scales::comma)+
  scale_x_continuous(labels = scales::dollar) +
  theme_bar+
  theme(legend.position = "bottom",
        text = element_text(size = 18))


girafe(ggobj = ext.time.distribution.plot, width_svg = 10, height_svg = 10) %>%
  girafe_options(opts_selection(type = "none"),
                 opts_sizing(rescale = FALSE))      

  

```


<br>

The table below provides each school district's total extended time revenue, the adjusted pupil units, the extended time pupil units, extended time revenue per extended time pupil units, and extended time revenue per adjusted pupil units.

As expected, every district either receives $0 in extended time revenue because they have no extended time pupil units, or $5,117 per extended time pupil units. 

However, the extended time revenue per adjusted pupil units varies significantly across districts. This indicates that schools have varying proportion of students that are participating in extended learning programs and/or offer a varying number of programs that qualify for funding.

The district table below shows that the districts receiving the most extended time revenue per APU vary in terms of being metro or rural districts. The highest is Dover-Eyota with $452 per APU. The other districts in the top are a mixture of metro and rural districts.

<br>

```{r extended time district}
extended.time.dist <- original %>%
  filter(group %in% c(1,4)) %>%
  select(`District Name`, dist, type, `County Name`, awadm22, exdayr, xwadm22) %>%
  mutate(`Extended time rev per ext time pupils` = exdayr / xwadm22,
         `Extended time rev per adjusted pupil units` = exdayr / awadm22) %>%
  rename(`District Number` = 2,
         `District Type` = 3,
         `Adjusted pupil units` = 5,
         `Total adjusted time revenue` = 6,
         `Adjusted extended time pupil units` = 7)

datatable(extended.time.dist, class = "cell-border stripe", filter = "top", rownames = FALSE,
          options = list(columnDefs = list(list(className = "dt-center", targets = 1:8)))) %>%
  formatCurrency(5, "", digits = 0) %>%
  formatCurrency(6, "$", digits = 0) %>%
  formatCurrency(7, "", digits = 0) %>%
  formatCurrency(8:9, "$", digits = 0)
```





 