---
title: "Scenarios"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 4
runtime: shiny
resource_files:
- Data/Shapefiles/County shapefiles/MNCounties_MNDOT.cpg
- Data/Shapefiles/County shapefiles/MNCounties_MNDOT.dbf
- Data/Shapefiles/County shapefiles/MNCounties_MNDOT.prj
- Data/Shapefiles/County shapefiles/MNCounties_MNDOT.sbn
- Data/Shapefiles/County shapefiles/MNCounties_MNDOT.sbx
- Data/Shapefiles/County shapefiles/MNCounties_MNDOT.shp.xml
- Data/Shapefiles/County shapefiles/MNCounties_MNDOT.shx
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(tidyverse)
library(sf)
library(ggrepel)
library(scales)
library(shiny)
library(shinycssloaders)
library(ggiraph)
library(kableExtra)
library(rmapshaper)
library(cowplot)
library(DT)
library(htmlwidgets)
library(RColorBrewer)
library(readxl)
library(janitor)
library(lubridate)
library(systemfonts)
reset_font_cache()
library(ggtext)
library(gmodels)

```

```{r loading jon docs and shapefiles, cache=TRUE, include=FALSE}
theme_bar <- theme_bw() +
  theme(panel.grid.major = element_line(color = "grey70", size = 0.1),
        panel.grid.minor = element_blank(),
        axis.ticks = element_blank(),
        axis.text = element_text(face = "bold"),
        panel.border = element_blank(),
        legend.background = element_rect(fill = "transparent", color = "transparent"),
        legend.key = element_rect(fill = "transparent"),
        legend.key.size = unit(1, "lines"),
        legend.margin = margin(0,0,0,0),
        legend.title = element_blank(),
        legend.text = element_text(margin = margin(l = 2)),
        text = element_text(family = "Arial"))

theme_line <- theme_bw() +
  theme(legend.background = element_rect(fill = "transparent", color = "transparent"),
        legend.key = element_rect(fill = "transparent"),
        legend.text = element_text(margin = margin(l = 2)),
        panel.grid.minor = element_blank(),
        panel.grid.major = element_line(color = "grey70", size = 0.1),
        axis.ticks = element_blank(),
        axis.text = element_text(face = "bold"),
        panel.border = element_blank(),
        legend.margin = margin(0,0,0,0),
        legend.key.size = unit(1, "lines"),
        text = element_text(family = "Arial"))


theme_sf <- theme_bw() +
  theme(axis.text.x=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks=element_blank(),
        panel.background = element_blank(),
        panel.grid.major = element_line(color = "white"),
        panel.border = element_blank(),
        plot.title = element_text(hjust = 0.5),
        legend.title = element_blank(),
        legend.text = element_text(margin = margin(l = 2)),
        legend.margin = margin(0,0,0,0),
        legend.key.size = unit(1, "lines"),
        text = element_text(family = "Arial"))

regions <- read_csv("Data/Join docs/county_regions.csv") %>%
    select(5,6) %>%
    unique() %>%
    mutate(edr = str_replace(edr, "  ", " "),
           planning.region = str_replace(planning.region, " Minnesota", ""),
           planning.region = fct_relevel(planning.region, "Northwest", "Northeast", "Central", "Seven County Mpls-St Paul", "Southwest", "Southeast"),
           edr = fct_relevel(edr, "EDR 1 - Northwest", "EDR 2 - Headwaters", "EDR 3 - Arrowhead", "EDR 4 - West Central", "EDR 5 - North Central", "EDR 6E- Southwest Central", "EDR 6W- Upper Minnesota Valley", "EDR 7E- East Central", "EDR 7W- Central", "EDR 8 - Southwest", "EDR 9 - South Central", "EDR 10 - Southeast", "EDR 11 - 7 County Twin Cities", "Minnesota"))

counties.regions <- read_csv("Data/Join docs/county_regions.csv") %>%
  rename(mif = `MIF Region`) %>%
  mutate(countyfp = formatC(countyfp, width = 3, flag = "0"),
         Name = str_to_title(Name),
         Name = str_replace(Name, "Q", "q"),
         Name = str_replace(Name, "Of The", "of the"),
         Name = str_replace(Name, "Mcleod", "McLeod"),
         planning.region = ifelse(Name == "Minnesota", "Minnesota", planning.region) ,
         edr = str_replace(edr, "  ", " "),
         planning.region = str_replace(planning.region, " Minnesota", ""),
         planning.region = fct_relevel(planning.region, "Northwest", "Northeast", "Central", "Seven County Mpls-St Paul", "Southwest", "Southeast"),
         edr = fct_relevel(edr, "EDR 1 - Northwest", "EDR 2 - Headwaters", "EDR 3 - Arrowhead", "EDR 4 - West Central", "EDR 5 - North Central", "EDR 6E- Southwest Central", "EDR 6W- Upper Minnesota Valley", "EDR 7E- East Central", "EDR 7W- Central", "EDR 8 - Southwest", "EDR 9 - South Central", "EDR 10 - Southeast", "EDR 11 - 7 County Twin Cities", "Minnesota"),
         mif = ifelse(is.na(mif), "TC", mif),
         mif = as.factor(mif),
         mif = fct_relevel(mif, "NW", "NE", "WC", "EC", "SW", "SE", "TC"))


color.edr <- c("Entirely rural" = "#009933", "Town/rural mix" = "#99CC33", "Urban/town/rural mix" = "#CC9966", "Entirely urban" = "#754C29", "Minnesota" = "black")

color.pr <- c("Northwest" = 	"#4575b4", "Northeast" = "grey", "Central" = "#fee090", "Seven County Mpls-St Paul" = "#d73027", "Southwest" = "#91bfdb", "Southeast" = "#fc8d59", "Minnesota" = "black")

color.edr <- c("EDR 1 - Northwest" = "#b3cde3", "EDR 2 - Headwaters" = "#8c96c6", "EDR 3 - Arrowhead" = "#fe9929", "EDR 4 - West Central" = "#8856a7", "EDR 5 - North Central" = "#810f7c", "EDR 6E- Southwest Central" = "#e5f5f9", "EDR 6W- Upper Minnesota Valley" = "#bdc9e1", "EDR 7E- East Central" = "#99d8c9", "EDR 7W- Central" = "#2ca25f", "EDR 8 - Southwest" = "#74a9cf", "EDR 9 - South Central" = "#0570b0", "EDR 10 - Southeast" = "#d7301f", "EDR 11 - 7 County Twin Cities" = "#d8b365", "Minnesota" = "black")

color.pr.edr <- c ("Northwest" = "#4575b4","Northeast" = "#e0f3f8", "Central" = "#fee090", "Seven County Mpls-St Paul" = "#d73027", "Southwest" = "#91bfdb", "Southeast" = "#fc8d59", "Minnesota" = "black", "EDR 1 - Northwest" = "#b3cde3", "EDR 2 - Headwaters" = "#8c96c6", "EDR 3 - Arrowhead" = "#fe9929", "EDR 4 - West Central" = "#8856a7", "EDR 5 - North Central" = "#810f7c", "EDR 6E- Southwest Central" = "#e5f5f9", "EDR 6W- Upper Minnesota Valley" = "#bdc9e1", "EDR 7E- East Central" = "#99d8c9", "EDR 7W- Central" = "#2ca25f", "EDR 8 - Southwest" = "#74a9cf", "EDR 9 - South Central" = "#0570b0", "EDR 10 - Southeast" = "#d7301f", "EDR 11 - 7 County Twin Cities" = "#d8b365")

mn_counties <- st_read("Data/Shapefiles/county shapefiles/MNCounties_MNDOT.shp", quiet = TRUE) %>%
  ms_simplify(keep = .01, keep_shapes = TRUE) %>%
  rename(countyfp = FIPS_CODE)

```

```{r master basic revenue}
original <- read_csv("Data/Formula/Master.csv") %>%
  mutate(Dem_Desc = fct_relevel(Dem_Desc, "Entirely rural", "Town/rural mix", "Urban/town/rural mix", "Entirely urban"),
         planning.region = fct_relevel(planning.region, "Northwest", "Northeast", "Central", "Seven County Mpls-St Paul", "Southwest", "Southeast"),
         edr = fct_relevel(edr, "EDR 1 - Northwest", "EDR 2 - Headwaters", "EDR 3 - Arrowhead", "EDR 4 - West Central", "EDR 5 - North Central", "EDR 6E- Southwest Central", "EDR 6W- Upper Minnesota Valley", "EDR 7E- East Central", "EDR 7W- Central", "EDR 8 - Southwest", "EDR 9 - South Central", "EDR 10 - Southeast", "EDR 11 - 7 County Twin Cities", "Minnesota"))

group.codes <- read_xlsx("Data/Formula/group codes.xlsx")

var.codes <- read_xlsx("Data/Formula/primary variable codes.xlsx")

```

<br>

First let's check to see what categories contribute the most revenue to school districts. Are there categories that provide more than another depending on the type of school, RUCA categories or regions?

First, let's lump all schools together and see what which categories provide the most. As the chart below shows, the basic education revenue provides BY FAR the most to school districts in Minnesota. Very far behind is compensatory which provides 6.4% of the total revenue given to schools.

<br>

```{r categories all school}
cat.all.schools <- original %>%
  filter(group %in% c(1,4)) %>%
  summarize(bsrev = sum(bsrev),
            exdayr = sum(exdayr),
            gtrev = sum(gtrev), 
            smlrev = sum(smlrev), 
            decrev = sum(decrev), 
            locaid = sum(locaid), 
            afdc22 = sum(afdc22), 
            totlep = sum(totlep), 
            totspar = sum(totspar), 
            cexaid = sum(cexaid, na.rm = TRUE), 
            tsparrev = sum(tsparrev), 
            eqaid = sum(eqaid), 
            hhaid = sum(hhaid), 
            penrev = sum(penrev)) %>%
  gather(key = "categories", value = "revenue", 1:14) %>%
  mutate(pct.revenue = revenue / sum(revenue),
         data_id = as.character(seq(n()))) %>%
  left_join(var.codes, by = c("categories" = "Code"))

cat.all.schools.plot <- ggplot(cat.all.schools, aes(fct_reorder(Category, pct.revenue), pct.revenue)) +
  geom_col_interactive(position = "dodge", aes(data_id = data_id, tooltip = paste(Category, "\nTotal revenue: ", dollar(revenue), "\nPercent of total revenue: ", percent(pct.revenue, accuracy = .1), sep = ""))) +
  geom_label(aes(label = percent(pct.revenue, accuracy = .1)), show.legend = FALSE, color = "black", size = 5, position = position_dodge(width = .9)) +
  coord_flip() +
  labs(x="", y = "", color="", title = "Percent share of total general education revenue\nformula")+
  scale_y_continuous(labels=scales::percent)+
  theme_bar+
  theme(legend.position = "none",
        text = element_text(size = 18))


girafe(ggobj = cat.all.schools.plot, width_svg = 10, height_svg = 10) %>%
  girafe_options(opts_selection(type = "none"),
                 opts_sizing(rescale = FALSE))      

```

<br>

Next, let's see if there are any big differences between public and charter schools.

<br>

```{r categories by group}
cat.group <- original %>%
  filter(group %in% c(1,4)) %>%
  group_by(group) %>%
  summarize(bsrev = sum(bsrev),
            exdayr = sum(exdayr),
            gtrev = sum(gtrev), 
            smlrev = sum(smlrev), 
            decrev = sum(decrev), 
            locaid = sum(locaid), 
            afdc22 = sum(afdc22), 
            totlep = sum(totlep), 
            totspar = sum(totspar), 
            cexaid = sum(cexaid, na.rm = TRUE), 
            tsparrev = sum(tsparrev), 
            eqaid = sum(eqaid), 
            hhaid = sum(hhaid), 
            penrev = sum(penrev)) %>%
  gather(key = "categories", value = "revenue", 2:15) %>%
  group_by(group) %>%
  mutate(pct.revenue = revenue / sum(revenue)) %>%
  ungroup() %>%
  mutate(data_id = as.character(seq(n()))) %>%
  left_join(var.codes, by = c("categories" = "Code")) %>%
  left_join(group.codes, by = c("group" = "Code")) %>%
  select(Definition, Category, revenue, pct.revenue, data_id)


cat.group.plot <- ggplot(cat.group, aes(fct_reorder(Category, pct.revenue), pct.revenue)) +
  facet_wrap(~Definition) +
  geom_col_interactive(position = "dodge", aes(data_id = data_id, tooltip = paste("School type: ", Definition, "n", Category, "\nTotal revenue: ", dollar(revenue), "\nPercent of total revenue: ", percent(pct.revenue, accuracy = .1), sep = ""))) +
  geom_label(aes(label = percent(pct.revenue, accuracy = .1)), show.legend = FALSE, color = "black", size = 5, position = position_dodge(width = .9)) +
  coord_flip() +
  labs(x="", y = "", color="", title = "Percent share of total general education revenue\nformula")+
  scale_y_continuous(labels=scales::percent)+
  theme_bar+
  theme(legend.position = "none",
        text = element_text(size = 18))


girafe(ggobj = cat.group.plot, width_svg = 10, height_svg = 10) %>%
  girafe_options(opts_selection(type = "none"),
                 opts_sizing(rescale = FALSE))      


```